\clearpage
\section{Scraping and Cleaning the Data}
\label{cleaning_data}
Data for this project was obtained from three different sources:
FightMetric.com, UFC.com and Wikipedia. Web scraping was used
in all cases by writing codes in Python which made use
of the Requests and BeautifulSoup modules. From FightMetric we scraped all the UFC fight
cards from UFC 1 to UFC 208. We also scraped the individual fighter
pages. From UFC.com we scraped the individual fighter pages which
give details on leg reach and education. We used to Wikipedia
to confirm or impute dates of birth, reach and height.

The FightMetric database of fighters has three pairs of fighters
with the same name: Michael McDonald, Dong Hyun Kim and
Tony Johnson. Each fighter name must be made unique since the name
serves as the primary key. One must also rename these fighters
in the FightMetric fight cards so that the two may be joined.
In the FightMetric fight cards there are two events with the same
name, namely, UFC Fight Night: Belfort vs Henderson. One should rename
one of the two events to group by event name. We also consolidated
event locations (e.g., ``Brooklyn, New York, USA'' became ``New York, New York, USA'').

The data obtained from UFC.com/fighter has duplicate names for
Gerard Gordeau and Dong Hyun Kim which were handled. Three fighters were
listed under the names ``... To be announced'', ``. To Be Determined'' and
``... To be determined'', which were removed. Some of the names had
two spaces between the first and last names. The fighter Matt Schnell
is listed as having a 72 inch leg reach which must be false.
After the cleaning there were 1897 fighters of which 533 were active. Note that
the UFC data set contains WEC fighters and maybe those from other promotions
acquired by the UFC.

The FightMetric and UFC reach data are in perfect agreement for 824 fighters.
They differ by 1 inch for 19 fighters and 2 inches for 18. Only 5 fighters
differ by more than 2 inches.
With respect to height, the two data sets agree perfectly for 1351 fighters,
differ by 1 inch for 131 fighters and 2 inches for 36. This leaves 19
cases where the height difference is greater than 2 inches. Dates of birth
are harder to check because the UFC only gives age.
Overall the agreement is quite good with maybe only 3\% of cases not
agreeing perfectly.
The author contacted FightMetric to ask about the discrepancies. Their reply
was that they believe their data to be the best source.

From UFC 1 to UFC 208 there have been 4096 fights between 1641 unique fighters. The
data obtained by scraping UFC.com is missing only 13 of these 1641. The missing
fighters are Benji Radach, Danillo Villefort, Eddie Mendez, Edilberto de Oliveira,
Joao Pierini, Julian Sanchez, Kit Cope, Luiz Cane, Masutatsu Yano, Nate Loughran,
Saeed Hosseini, Scott Smith and Tito Ortiz. Note that fuzzy matching was useful
in matching many of the names between the FightMetric and UFC data sets.

