\documentclass[12pt]{article}

\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{etoolbox}
\usepackage{dcolumn}
\usepackage{hyperref}

\BeforeBeginEnvironment{tabular}{\begin{center}\footnotesize}
\AfterEndEnvironment{tabular}{\end{center}}

\title{\Large\bfseries Statistical Analysis and Predictive Modeling of Mixed Martial Arts Fighting: From UFC~1~to~UFC~208}
\author{\large Jonathan D. Halverson, Ph.D.\\ \tiny halverson.jonathan at gmail}
\date{\large \today}

\usepackage[export]{adjustbox}
\usepackage{amsmath}
\usepackage[in]{fullpage}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage[font=footnotesize,margin=2cm]{caption}
\usepackage{indentfirst}

\renewcommand{\thesection}{\Roman{section}.} 
\renewcommand{\thesubsection}{\thesection.\Roman{subsection}}
\usepackage{titlesec}
\titleformat*{\section}{\large\bfseries}
\titleformat*{\subsection}{\normalsize\bfseries}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
%\cfoot{\tiny Paul Valenti Special Edition}
\rfoot{\thepage}

\begin{document}
\maketitle
\thispagestyle{fancy}
\noindent
\textbf{Summary} -- Here we analyze UFC data from 1993 through February 2017.
We find southpaws have a significant advantage against
orthodox fighters, ring rust takes effect around 9-12 months, and fighters tend to lose
more than they win after around age 33 with fighter performance on the decline after about age 28.
Average age, height and reach has increased in almost all weight classes
with the average welterweight standing two inches taller today than only a decade ago.
Quite remarkably these values are still increasing.
We find that college graduates beat non-graduates
56.4 percent of the time.
In the last section, we utilize
a branch of Artificial Intelligence called machine learning to
train models such as artificial neural networks and random forests to
predict fight outcomes. Our best model has an accuracy of approximately 60\%. We argue that
prediction is made difficult by the lack of fighters with UFC experience. 
That is, in 78\% of all UFC bouts one or both fighters had only four or fewer previous UFC fights.

\input{intro/intro.tex}
\input{age/age.tex}
\input{reach_height/reach_height.tex}
\input{southpaw/southpaw.tex}
\input{offense_defense/offense_defense.tex}
\input{finishes/finishes.tex}
\input{education/education.tex}
\input{prediction/prediction.tex}
\input{conclusions/conclusions.tex}

\section*{Acknowledgements}
We are grateful to FightMetric for answering our questions about differences
in the date of birth, reach and height between different data sources.


\appendix
\input{cleaning/cleaning.tex}

\section{UFC Fighters with College Experience}
\label{appendix_education}
The data in the table below was obtained from \url{ufc.com/fighter} on February 28, 2017.
All fighters in the UFC database with a listing in either College or Degree are shown.
\begin{center}
\input{education/education_sample.tex}
\end{center}

%\clearpage
%\section{Every UFC Fight up to UFC 208}
%\label{every_ufc_fight}
%\begin{center}
%\input{../fights_table.tex}
%\end{center}

\end{document}
